Uncertainty propagation in neural networks for sparse coding

### 2018-05-13 paper/BayesianLISTA.tex
paper: nips style and attempts to shorten the text

\documentclass[letterpaper]{article}
\usepackage{proceed2e}
\usepackage[margin=1in]{geometry}
\usepackage{mathtools}
\usepackage{xfrac}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{subfig}
\usepackage{hyperref}
\usepackage{xcolor}

@@

\usepackage{natbib}

@@

\section{\uppercase{Introduction}}
Though the idea of Bayesian learning in neural networks is not new \citep{neal2012bayesian}, it has regained its attention recently with the development of distributed approximate inference techniques \citep{li2015stochastic, hoffman2013stochastic}  and general boost in popularity of deep learning. In some spheres, such as self-driving cars or healthcare, uncertainty estimates are very important and Bayesian learning allows to obtain them.

@@

The main contributions of this paper can be summarised~as:
\begin{itemize}
\item for the first time we propose a method for uncertainty propagation through the soft thresholding nonlinearity for a Bayesian neural network;
\item we develop an efficient posterior inference algorithm for weights and outputs of neural networks with the soft thresholding nonlinearity;
\item we design a novel Bayesian neural network for sparse coding.
\end{itemize}

The paper is organised as following: first, we review neural networks for sparse coding in section \ref{sec:nn_sc} and present a novel Bayesian neural network in section \ref{sec:bayesian_lista}. We then describe uncertainty propagation to obtain the spike and slab distribution for an output of a network in section \ref{sec:fprop} and how probabilistic backpropagation is organised to update distributions of weights in section \ref{sec:backpropagation}. After that, we show the experimental results of the algorithm in section \ref{sec:experiments} and finish with discussion and conclusions in sections \ref{sec:discussion} and \ref{sec:conclusions}.

\section{\uppercase{Neural networks for sparse coding}}
\label{sec:nn_sc}
The sparse coding problem is the linear regression with the assumption of sparse weights
\begin{equation}
\label{eq:regression_problem}
\mathbf{y} = \mathbf{X}\boldsymbol\beta + \varepsilon
\end{equation}
where $\mathbf{y} \in \mathbb{R}^K$ is the observation, $\mathbf{X} \in \mathbb{R}^{K \times D}$ is the design matrix, $\boldsymbol\beta \in \mathbb{R}^D$ is the unknown vector of weights with many elements $\beta_d$ equal to zero, $\varepsilon \in \mathbb{R}^K$ is the independent Gaussian noise.

One of the popular families of the sparse coding algorithms is the iterative shrinkage algorithms, and notably the Iterative Shrinkage and Thresholding Algorithm (\textsc{ista}) \citep{daubechies2004iterative}. It iteratively obtains the new approximation of the coefficient vector $\widehat{\boldsymbol\beta}_l$ at the iteration $l$ as the linear transformation of the input $\mathbf{y}$ with the previous approximation $\widehat{\boldsymbol\beta}_{l-1}$ and then propagates the new approximation through the soft thresholding function $h_\lambda(\cdot)$
\begin{equation}
h_\lambda(v) = \text{sgn}(v) \max(|v| - \lambda, 0)
\end{equation}
where $\lambda$ is a shrinkage parameter.
The linear transformation is
\begin{equation}
\widehat{\boldsymbol\beta}_l = \mathbf{W}\mathbf{y} + \mathbf{S}_l\widehat{\boldsymbol\beta}_{l-1}
\end{equation}
with weights $\mathbf{W} = \mathbf{X}^\top / E$, $E$ - upper bound of the largest eigenvalue of $\mathbf{X}^\top\mathbf{X}$ and $\mathbf{S} = \mathbf{I}_{D \times D} - \mathbf{W}\mathbf{X}$ ($\mathbf{I}_{D \times D}$ is the identity matrix of size $D$).

@@

Matrices $\mathbf{W}$, $\mathbf{S}$ are the parameters that are initialised as in the \textsc{ista} and then updated with the backpropagation algorithm. Vectors $\mathbf{c}_l$, $\mathbf{b}$ are intermediate vectors that describe forward propagation.

\section{\uppercase{Bayesian neural network for sparse coding}}
\label{sec:bayesian_lista}

This section presents the proposed Bayesian neural network for sparse coding. To formulate the Bayesian version of the \textsc{lista} we impose the prior distributions on the unknown weights
\begin{subequations}
\label{eq:ws}
\begin{align}
&p(\mathbf{W}) = \prod_{d=1}^D\prod_{k=1}^K \mathcal{N}(w_{ij} | 0, \eta^{-1}), \\
&p(\mathbf{S}) = \prod_{d'=1}^D\prod_{d''=1}^D \mathcal{N}(s_{d'd''} | 0, \eta^{-1}),
\end{align}
\end{subequations}
where $w_{ij}$ is a component of the matrix $\mathbf{W}$, $s_{d'd''}$ is a component of the matrix $\mathbf{S}$, $\eta$ is the precision of the Gaussian distribution.
The likelihood of $\mathbf{B}$ is defined as
\begin{align}
p(&\mathbf{B}| \mathbf{Y}, \mathbf{W}, \mathbf{S}, \gamma, \lambda) \nonumber \\
\label{eq:likelihood}
= &\prod_{n=1}^N\prod_{d=1}^D\mathcal{N}(\beta_d^{(n)}; [f(\mathbf{y^{(n)}}; \mathbf{W}, \mathbf{S}, \lambda)]_d, \gamma^{-1}),
\end{align}
where $[\cdot]_d$ denotes the d-th component of a vector, $f(\mathbf{y}^{(n)}; \mathbf{W}, \mathbf{S}, \lambda)$ is the output of the Bayesian \textsc{lista} neural network, $\gamma$ is the precision of the Gaussian distribution.
The prior of the introduced Gaussian precisions are set to the Gamma distribution with parameters~$a^{\cdot}$ and~$b^{\cdot}$:
\begin{subequations}
\label{eq:gamma_eta}
\begin{align}
&p(\gamma) = \text{Gam}\left(\gamma; a^{\gamma}, b^{\gamma}\right), \\
&p(\eta) = \text{Gam}\left(\eta; 	a^{\eta}, b^{\eta}\right)
\end{align}
\end{subequations}

The posterior distribution is then
\begin{align}
p(&\mathbf{W}, \mathbf{S}, \gamma, \eta | \mathbf{B}, \mathbf{Y}, \lambda) \nonumber \\
\label{eq:posterior}
= & \frac{p(\mathbf{B} | \mathbf{Y}, \mathbf{W},  \mathbf{S}, \gamma, \lambda) p(\mathbf{W} | \eta )p(\mathbf{S} | \eta) p(\eta) p(\gamma)}{p(\mathbf{B} | \mathbf{Y}, \lambda)}
\end{align}
The shrinkage parameter $\lambda$ is a hyperparameter of the model.

\section{\uppercase{Uncertainty propagation through soft thresholding}}
\label{sec:fprop}
For every layer of the Bayesian \textsc{lista} we assume that $\widehat{\boldsymbol\beta}_{l-1}$ has the spike and slab distribution with parameters: $\boldsymbol\omega$ is a probability of a spike, $\mathbf{m}$ is a mean of a slab Gaussian distribution, and $\mathbf{v}$ is a variance of the slab distribution:
\begin{equation}
[\widehat{\boldsymbol\beta}_{l-1}]_d \sim \omega_d \delta_0 + (1 - \omega_d)\mathcal{N}(m_d, v_d),
\end{equation}
where $\omega_d$, $m_d$, and $v_d$ are the components of vectors $\boldsymbol\omega$, $\mathbf{m}$, and $\mathbf{v}$, respectively, $\delta_0$ is the delta-function that represents a spike.

@@

\subsection{\uppercase{Initialisation dense layer}}
The first step in the Bayesian \textsc{lista} is (\ref{eq:first_layer}). The matrix $\mathbf{W}$ consists of Gaussian-distributed components: $w_{dk} \sim \mathcal{N}(m^w_{dk}, v^w_{dk})$, and $y$ is a deterministic vector. Since the sum of independent Gaussian-distributed variables is a Gaussian variable~\citep{eisenberg2008sum}, $\mathbf{b}$ is a vector of Gaussian-distributed components $b_d \sim \mathcal{N}(m^b_d, v^b_d)$
\begin{equation}
\label{eq:matrix_vector_product}
m^b_d = \sum_{k=1}^Ky_k m^w_{dk}, \qquad
v^b_d = \sum_{k=1}^Ky_k^2v^w_{dk}.
\end{equation}

\subsection{\uppercase{Soft thresholding nonlinearity}}
At the second step (\ref{eq:thr_first}) of the Bayesian \textsc{lista} the Gaussian vector $\mathbf{b}$ is taken as an input of the soft thresholding function.

@@

The proof of Lemma~\ref{thm:soft_thresholding} and expressions for $\omega^*$, $m^*$ and $v^*$ are given in the supplementary materials. Based on the results of Lemma~\ref{thm:soft_thresholding} we approximate the distribution of $\widehat{\boldsymbol\beta}_0$ from (\ref{eq:thr_first}) with a spike and slab distribution with parameters derived in the lemma: the spike probability $\omega$ is equal to $\omega^*$, the slab mean $m$ and variance $v$ are set to the corresponding parameters of the truncated Gaussian $m^*$ and $v^*$.

\subsection{\uppercase{Main layers}}
The distributions of the inputs at the step (\ref{eq:l_dense_layer}) of the Bayesian \textsc{lista} for each $l$ are: the vector $\mathbf{b}$ and matrix $\mathbf{S}$ that consist of Gaussian-distributed components and $\widehat{\boldsymbol\beta}_{l-1}$, which is a vector of the spike and slab random variables.

In order to determine the distribution of the output $\mathbf{c}_l$ we first formulate two lemmas.

\begin{lemma}[Moments of a spike and slab distribution]
\label{thm:moments_spsl}
Let a random variable $\xi$ have a spike and slab distribution with probability of spike $\omega$, slab mean $m$ and slab variance $v$. Then its moments are
\begin{subequations}
\begin{align}
\mathbb{E}\xi &= (1-\omega)m \\
\operatorname{Var}\xi & = (1-\omega)(v + \omega m^2)
\end{align}
\end{subequations}
\end{lemma}

\begin{lemma}[Product of a Gaussian matrix and a spike and slab vector]
  \label{thm:matrix_vector}
Let $\mathbf{S} \in \mathbb{R}^{D \times D}$ be a matrix of independent Gaussian-distributed random variables: $s_{d'd''} \sim \mathcal{N}(m^s_{d'd''}, v^s_{d'd''})$, and $\widehat{\boldsymbol\beta }\in \mathbb{R}^D$ be a vector with spike-and-slab distributed variables: $\widehat{\beta}_d \sim \omega_d \delta_0 + (1 - \omega_d)\mathcal{N}(m_d, v_d)$. The components of the matrix $\mathbf{S}$ and the vector $\widehat{\boldsymbol\beta}$ are mutual independent. Let $\mathbf{e} \in \mathbb{R}^{D}$ denote the product $\mathbf{S} \widehat{\boldsymbol\beta}$. Then the marginal mean and variance of elements $e_d$ of the vector $\mathbf{e}$ are:
\begin{subequations}
\begin{align}
 \mathbb{E}e_d &= \sum_{d'=1}^D m^s_{dd'}(1-\omega_{d'})m_{d'}, \\
 \begin{split}
 \operatorname{Var}e_d &= \sum_{d'=1}^D [(m^s_{dd'})^2(1-\omega_{d'})^2v_{d'} \\
 & {}+ (1-\omega_{d'})^2(m_{d'})^2v^s_{dd'} + v^s_{dd'}(1-\omega_{d'})^2v_{d'}].
 \end{split}
 \end{align}
\end{subequations}
 \end{lemma}

@@

Let $\mathbf{e}_l$ denote the product $\mathbf{S}\widehat{\boldsymbol\beta}_{l-1}$ at the step (\ref{eq:l_dense_layer}) of the Bayesian \textsc{lista}. We assume that $\mathbf{S}$ and $\widehat{\boldsymbol\beta}_{l-1}$ are mutually independent. Then according to the Central Limit Theorem $[ \mathbf{e}_l ]_d$ can be approximated as a Gaussian-distributed variable when $D$ is sufficiently large. The parameters of the of the approximating Gaussian distribution are set to the marginal mean and variance given in Lemma~\ref{thm:matrix_vector}. The quality of this approximation is discussed in section \ref{sec:approx_quality}.

@@

\begin{figure}[!t]
\subfloat[Approximation of the \newline product of Gaussians]{\includegraphics[width=0.5\columnwidth]{d_testing}\label{fig:d_testing}}
%\label{fig:d_testing}
\subfloat[Approximation of the propagation through soft thresholding]{\includegraphics[width=0.5\columnwidth]{z_new_testing}\label{fig:z_new_testing}}
\caption{Approximations of the Bayesian \textsc{lista}}
%\label{fig:z_new_testing}
\end{figure}

\subsection{\uppercase{Bayesian lista forward propagation}}
We have covered all steps of the Bayesian \textsc{lista} and derived distributions for outputs of these steps. Now we can formulate the main result of this section that describes how the proposed uncertainty propagation works.

@@

\subsection{\uppercase{Approximation quality}}
\label{sec:approx_quality}
We have used two approximations in forward propagation of uncertainty. First, at the step (\ref{eq:l_dense_layer}) during the computation of $\mathbf{e}_l$, a Gaussian matrix is multiplied by a spike and slab vector and their product is approximated with the Gaussian distribution. Second, at the steps (\ref{eq:thr_first}) and (\ref{eq:l_thr}) the result of soft thresholding of a Gaussian vector is approximated with the spike and slab distribution. In this section we demonstrate that these approximations are close to the real distributions.

Figure \ref{fig:d_testing} demonstrates the comparison of the sampled distribution and approximated distribution for Lemma \ref{thm:matrix_vector}. For the sampled distribution, $10000$ values are sampled from the Gaussian matrix and the spike and slab vector and their product is computed, then one of the dimensionalities is plotted. The parameters of the approximated distribution is computed according to Lemma \ref{thm:matrix_vector}.

Figure \ref{fig:z_new_testing} demonstrates the comparison of the sampled distribution and approximated distribution for Lemma \ref{thm:soft_thresholding}. For the sampled distribution, $10000$ values are sampled from the Gaussian vector and propagated through soft thresholding, then one of the dimensionalities is plotted. The parameters of the approximated distribution is computed according to Lemma \ref{thm:soft_thresholding}.

@@

\section{\uppercase{Backpropagation}}
\label{sec:backpropagation}

The exact posterior (\ref{eq:posterior}) is approximated with a factorised distribution
\begin{align}
\label{eq:approximating_dsitribution}
\begin{split}
q(\mathbf{W}, \mathbf{S}, \gamma, \eta) &= \prod_{d=1}^D\prod_{k=1}^K \mathcal{N}(w_{dk} | m^w_{dk}, v^w_{dk}) \\
&\times \prod_{d'=1}^D\prod_{d''=1}^D \mathcal{N}(s_{d'd''} | m^s_{d'd''}, v^s_{d'd''}) \\
&\times \text{Gam}(\gamma; a^\gamma, b^\gamma) \text{Gam}(\eta; a^\eta, b^\eta)
\end{split}
\end{align}

@@

When a factor from $p$ is incorporated into $q$, $q$ as a function of Gaussian-distributed weights $\mathbf{W}$ and $\mathbf{S}$ has the following form:
\begin{equation}
q(a) = Z^{-1}f(a)\mathcal{N}(a; m, v)
\end{equation}
where $Z$ is the normalisation constant and $f(a)$ is an arbitrary function, $a \in \{w_{dk}, s_{d'd''}; \forall d, k, d', d''\}$.

According to \cite{minka2001thesis}, new parameters of the Gaussian distribution for $a$ can be computed as
\begin{align}
m^{\text{new}} &= m + v \frac{\partial \log Z}{\partial m}, \\
v^{\text{new}} &= v - v^2\left[ \left(\frac{\partial \log Z}{\partial m}\right)^2 - 2 \frac{\partial \log Z}{\partial v}\right]
\end{align}

@@

\subsection{\uppercase{Likelihood}}

>>>>

\documentclass{article}
\usepackage{mathtools}
\usepackage{xfrac}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{subfig}
\usepackage{hyperref}
\usepackage{xcolor}

@@

\usepackage{natbib}
\usepackage{nips_2018}

@@

\section{Introduction}
Though the idea of Bayesian learning in neural networks is not new \citep{neal2012bayesian}, it has regained its attention recently with the development of distributed approximate inference techniques \citep{li2015stochastic, hoffman2013stochastic}  and general boost in popularity of deep learning. In some spheres, such as self-driving cars or healthcare, uncertainty estimates are very important and Bayesian learning allows to obtain them.

@@

The main contributions of this paper are: (\textit{i}) for the first time we propose a method for uncertainty propagation through the soft thresholding nonlinearity for a Bayesian neural network; (\textit{ii}) we develop an efficient posterior inference algorithm for weights and outputs of neural networks with the soft thresholding nonlinearity; (\textit{iii}) we design a novel Bayesian neural network for sparse coding.

%The paper is organised as following: first, we review neural networks for sparse coding in section \ref{sec:nn_sc} and present a novel Bayesian neural network in section \ref{sec:bayesian_lista}. We then describe uncertainty propagation to obtain the spike and slab distribution for an output of a network in section \ref{sec:fprop} and how probabilistic backpropagation is organised to update distributions of weights in section \ref{sec:backpropagation}. After that, we show the experimental results of the algorithm in section \ref{sec:experiments} and finish with discussion and conclusions in sections \ref{sec:discussion} and \ref{sec:conclusions}.

\section{Neural networks for sparse coding}
\label{sec:nn_sc}
The sparse coding problem is the linear regression with the assumption of sparse weights
\begin{equation}
\label{eq:regression_problem}
\mathbf{y} = \mathbf{X}\boldsymbol\beta + \varepsilon
\end{equation}
where $\mathbf{y} \in \mathbb{R}^K$ is the observation, $\mathbf{X} \in \mathbb{R}^{K \times D}$ is the design matrix, $\boldsymbol\beta \in \mathbb{R}^D$ is the unknown vector of weights with many elements $\beta_d$ equal to zero, $\varepsilon \in \mathbb{R}^K$ is the independent Gaussian noise.

One of the popular families of the sparse coding algorithms is the iterative shrinkage algorithms, and notably the Iterative Shrinkage and Thresholding Algorithm (\textsc{ista}) \citep{daubechies2004iterative}. It iteratively obtains the new approximation of the coefficient vector $\widehat{\boldsymbol\beta}_l$ at the iteration $l$ as the linear transformation of the input $\mathbf{y}$ with the previous approximation $\widehat{\boldsymbol\beta}_{l-1}$ and then propagates the new approximation through the soft thresholding function $h_\lambda(\cdot)$
\begin{equation}
h_\lambda(v) = \text{sgn}(v) \max(|v| - \lambda, 0)
\end{equation}
where $\lambda$ is a shrinkage parameter.
The linear transformation is
\begin{equation}
\widehat{\boldsymbol\beta}_l = \mathbf{W}\mathbf{y} + \mathbf{S}\widehat{\boldsymbol\beta}_{l-1}
\end{equation}
with weights $\mathbf{W} = \mathbf{X}^\top / E$, $E$ - upper bound of the largest eigenvalue of $\mathbf{X}^\top\mathbf{X}$ and $\mathbf{S} = \mathbf{I}_{D \times D} - \mathbf{W}\mathbf{X}$ ($\mathbf{I}_{D \times D}$ is the identity matrix of size $D$).

@@

Matrices $\mathbf{W}$, $\mathbf{S}$ are the parameters that are initialised as in the \textsc{ista} and then updated with the backpropagation algorithm. Vectors $\mathbf{c}_l$, $\mathbf{b}$ are auxiliary vectors that describe forward propagation.

\section{Bayesian neural network for sparse coding}
\label{sec:bayesian_lista}

This section presents the proposed Bayesian neural network for sparse coding. To formulate the Bayesian version of the \textsc{lista} we impose the prior distributions on the unknown weights
\begin{equation}
\label{eq:ws}
p(\mathbf{W}) = \prod_{d=1}^D\prod_{k=1}^K \mathcal{N}(w_{ij} | 0, \eta^{-1}), \qquad
p(\mathbf{S}) = \prod_{d'=1}^D\prod_{d''=1}^D \mathcal{N}(s_{d'd''} | 0, \eta^{-1}),
\end{equation}
where $w_{ij}$ is a component of the matrix $\mathbf{W}$, $s_{d'd''}$ is a component of the matrix $\mathbf{S}$, $\eta$ is the precision of the Gaussian distribution.
The likelihood of $\mathbf{B}$ is defined as
\begin{equation}
\label{eq:likelihood}
p(\mathbf{B}| \mathbf{Y}, \mathbf{W}, \mathbf{S}, \gamma, \lambda) = \prod_{n=1}^N\prod_{d=1}^D\mathcal{N}(\beta_d^{(n)}; [f(\mathbf{y^{(n)}}; \mathbf{W}, \mathbf{S}, \lambda)]_d, \gamma^{-1}),
\end{equation}
where $[\cdot]_d$ denotes the d-th component of a vector, $f(\mathbf{y}^{(n)}; \mathbf{W}, \mathbf{S}, \lambda)$ is the output of the Bayesian \textsc{lista} neural network, $\gamma$ is the precision of the Gaussian distribution.
The prior of the introduced Gaussian precisions are set to the Gamma distribution with parameters~$a^{\cdot}$ and~$b^{\cdot}$:
\begin{equation}
\label{eq:gamma_eta}
p(\gamma) = \text{Gam}\left(\gamma; a^{\gamma}, b^{\gamma}\right), \qquad
p(\eta) = \text{Gam}\left(\eta; 	a^{\eta}, b^{\eta}\right)
\end{equation}

The posterior distribution is then
\begin{equation}
\label{eq:posterior}
p(\mathbf{W}, \mathbf{S}, \gamma, \eta | \mathbf{B}, \mathbf{Y}, \lambda) = \frac{p(\mathbf{B} | \mathbf{Y}, \mathbf{W},  \mathbf{S}, \gamma, \lambda) p(\mathbf{W} | \eta )p(\mathbf{S} | \eta) p(\eta) p(\gamma)}{p(\mathbf{B} | \mathbf{Y}, \lambda)}
\end{equation}
The shrinkage parameter $\lambda$ is a hyperparameter of the model.

\section{Uncertainty propagation through soft thresholding}
\label{sec:fprop}
For every layer of the Bayesian \textsc{lista} we assume that $\widehat{\boldsymbol\beta}_{l-1}$ has the spike and slab distribution with parameters: $\boldsymbol\omega$ is a probability of a spike, $\mathbf{m}$ is a mean of a slab Gaussian distribution, and $\mathbf{v}$ is a variance of the slab distribution:
\begin{equation}
[\widehat{\boldsymbol\beta}_{l-1}]_d \sim \omega_d \delta_0 + (1 - \omega_d)\mathcal{N}(m_d, v_d),
\end{equation}
where $\omega_d$, $m_d$, and $v_d$ are the components of vectors $\boldsymbol\omega$, $\mathbf{m}$, and $\mathbf{v}$, respectively, $\delta_0$ is the delta-function that represents a spike.

@@

\subsection{Initialisation dense layer}
The first step in the Bayesian \textsc{lista} is (\ref{eq:first_layer}). The matrix $\mathbf{W}$ consists of Gaussian-distributed components: $w_{dk} \sim \mathcal{N}(m^w_{dk}, v^w_{dk})$, and $y$ is a deterministic vector. Since the sum of independent Gaussian-distributed variables is a Gaussian variable~\citep{eisenberg2008sum}, $\mathbf{b}$ is a vector of Gaussian-distributed components $b_d \sim \mathcal{N}(m^b_d, v^b_d)$
\begin{equation}
\label{eq:matrix_vector_product}
m^b_d = \sum_{k=1}^Ky_k m^w_{dk}, \qquad
v^b_d = \sum_{k=1}^Ky_k^2v^w_{dk}.
\end{equation}

\subsection{Soft thresholding nonlinearity}
At the second step (\ref{eq:thr_first}) of the Bayesian \textsc{lista} the Gaussian vector $\mathbf{b}$ is taken as an input of the soft thresholding function.

@@

The proof of Lemma~\ref{thm:soft_thresholding} and expressions for $\omega^*$, $m^*$ and $v^*$ are given in the supplementary materials. Based on the results of Lemma~\ref{thm:soft_thresholding} we approximate the distribution of $\widehat{\boldsymbol\beta}_0$ from (\ref{eq:thr_first}) with a spike and slab distribution with parameters derived in the lemma: the spike probability $\omega$ is equal to $\omega^*$, the slab mean $m$ and variance $v$ are set to the corresponding parameters of the truncated Gaussian $m^*$ and $v^*$. The quality of this approximation is discussed in the supplementary materials.

\subsection{Main layers}
The distributions of the inputs at the step (\ref{eq:l_dense_layer}) of the Bayesian \textsc{lista} for each $l$ are: the vector $\mathbf{b}$ and matrix $\mathbf{S}$ that consist of Gaussian-distributed components and $\widehat{\boldsymbol\beta}_{l-1}$, which is a vector of the spike and slab random variables. In order to determine the distribution of the output $\mathbf{c}_l$ we first formulate two lemmas.

\begin{lemma}[Moments of a spike and slab distribution]
\label{thm:moments_spsl}
Let a random variable $\xi$ have a spike and slab distribution with probability of spike $\omega$, slab mean $m$ and slab variance $v$. Then its moments are
\begin{equation}
\mathbb{E}\xi = (1-\omega)m, \qquad \operatorname{Var}\xi = (1-\omega)(v + \omega m^2)
\end{equation}
\end{lemma}

\begin{lemma}[Product of a Gaussian matrix and a spike and slab vector]
  \label{thm:matrix_vector}
Let $\mathbf{S} \in \mathbb{R}^{D \times D}$ be a matrix of independent Gaussian-distributed random variables: $s_{d'd''} \sim \mathcal{N}(m^s_{d'd''}, v^s_{d'd''})$, and $\widehat{\boldsymbol\beta }\in \mathbb{R}^D$ be a vector with spike-and-slab distributed variables: $\widehat{\beta}_d \sim \omega_d \delta_0 + (1 - \omega_d)\mathcal{N}(m_d, v_d)$. The components of the matrix $\mathbf{S}$ and the vector $\widehat{\boldsymbol\beta}$ are mutual independent. Let $\mathbf{e} \in \mathbb{R}^{D}$ denote the product $\mathbf{S} \widehat{\boldsymbol\beta}$. Then the marginal mean and variance of elements $e_d$ of the vector $\mathbf{e}$ are:
\begin{subequations}
\begin{align}
 \mathbb{E}e_d &= \sum_{d'=1}^D m^s_{dd'}(1-\omega_{d'})m_{d'}, \\
 \begin{split}
 \operatorname{Var}e_d &= \sum_{d'=1}^D [(m^s_{dd'})^2(1-\omega_{d'})^2v_{d'} + (1-\omega_{d'})^2(m_{d'})^2v^s_{dd'} + v^s_{dd'}(1-\omega_{d'})^2v_{d'}].
 \end{split}
 \end{align}
\end{subequations}
 \end{lemma}

@@

Let $\mathbf{e}_l$ denote the product $\mathbf{S}\widehat{\boldsymbol\beta}_{l-1}$ at the step (\ref{eq:l_dense_layer}) of the Bayesian \textsc{lista}. We assume that $\mathbf{S}$ and $\widehat{\boldsymbol\beta}_{l-1}$ are mutually independent. Then according to the Central Limit Theorem $[ \mathbf{e}_l ]_d$ can be approximated as a Gaussian-distributed variable when $D$ is sufficiently large. The parameters of the of the approximating Gaussian distribution are set to the marginal mean and variance given in Lemma~\ref{thm:matrix_vector}. The supplementary materials provide a discussion of the quality of this approximation.

@@

\subsection{Bayesian \textsc{lista} forward propagation}
We have covered all steps of the Bayesian \textsc{lista} and derived distributions for outputs of these steps. Now we can formulate the main result of this section that describes how the proposed uncertainty propagation works.

@@

@@

\section{Backpropagation}
\label{sec:backpropagation}

The exact posterior (\ref{eq:posterior}) is approximated with a factorised distribution
\begin{align}
\label{eq:approximating_dsitribution}
\begin{split}
q(\mathbf{W}, \mathbf{S}, \gamma, \eta) &= \prod_{d=1}^D\prod_{k=1}^K \mathcal{N}(w_{dk} | m^w_{dk}, v^w_{dk}) \prod_{d'=1}^D\prod_{d''=1}^D \mathcal{N}(s_{d'd''} | m^s_{d'd''}, v^s_{d'd''}) \\
&\times \text{Gam}(\gamma; a^\gamma, b^\gamma) \text{Gam}(\eta; a^\eta, b^\eta)
\end{split}
\end{align}

@@

When a factor from $p$ is incorporated into $q$, $q$ as a function of Gaussian-distributed weights $\mathbf{W}$ and $\mathbf{S}$ has the following form:
\begin{equation}
q(a) = Z^{-1}f(a)\mathcal{N}(a; m, v)
\end{equation}
where $Z$ is the normalisation constant and $f(a)$ is an arbitrary function, $a \in \{w_{dk}, s_{d'd''}\}$.

According to \cite{minka2001thesis}, new parameters of the Gaussian distribution for $a$ can be computed as
\begin{equation}
m^{\text{new}} = m + v \frac{\partial \log Z}{\partial m}, \qquad
v^{\text{new}} = v - v^2\left[ \left(\frac{\partial \log Z}{\partial m}\right)^2 - 2 \frac{\partial \log Z}{\partial v}\right]
\end{equation}

@@

With the likelihood factors~(\ref{eq:likelihood}) of the true posterior we employ the ADF approach and iteratively incorporate them into the approximating distribution $q$. The normalisation constant of the approximating distribution $q$ with the likelihood term for the data point $n$ incorporated can be computed as follows (to simplify notation the superscript $(n)$ is omitted)
\begin{align}
Z & = \int \prod_{d=1}^{D} \mathcal{N}(\beta_d | f(\mathbf{y} ; \mathbf{S}, \mathbf{W}, \lambda), \gamma^{-1}) q(\mathbf{W}, \mathbf{S}, \gamma, \eta) \mathrm{d}\mathbf{W} \mathrm{d}\mathbf{S} \mathrm{d}\gamma \mathrm{d}\eta \nonumber\\
\label{eq:Z}
& \approx \prod_{d=1}^D \left[\omega^{\widehat{\boldsymbol\beta}}_d  \mathcal{T}\left(\beta_d | 0, \beta^\gamma / \alpha^\gamma, 2\alpha^\gamma\right) + \vphantom{m^{\widehat{\boldsymbol\beta}}_d} \left(1 - \omega^{\widehat{\boldsymbol\beta}}_d\right)\mathcal{N}\left(\beta_d | m^{\widehat{\boldsymbol\beta}}_d,  \beta^\gamma / (\alpha^\gamma - 1) + v^{\widehat{\boldsymbol\beta}}\right)\right].
\end{align}
Parameters of the approximating posterior distribution are then updated with the derivatives of this normalisation constant.

Prior factors from $p$ (\ref{eq:ws}), (\ref{eq:gamma_eta}) are incorporated into $q$ with the EP algorithm~\citep{hernandez2015probabilistic}, i.e. they replace the corresponding approximating factors from $q$ and then $q$ is updated to minimise its KL divergence with $q$ that has the true factor incorporated.

\section{Experiments}
\label{sec:experiments}
The proposed Bayesian \textsc{lista} is evaluated in the context of the sparse coding problem with an overcomplete dictionary, where the number of measurements $K$ is much smaller than the dimensionality of the vector $\boldsymbol\beta$.

We compare the proposed Bayesian \textsc{lista} with the classical \textsc{lista}~\citep{gregor2010learning} in terms of the predictive accuracy. As baselines we also use \textsc{ista} \citep{daubechies2004iterative} and Fast \textsc{ista} (\textsc{fista}) \citep{beck2009fast}. \textsc{fista} adds the momentum to \textsc{ista} and improves its convergence speed. We set the number of iterations in these algorithms and the number of layers in the Bayesian and classical \textsc{lista} networks to $L$. To measure the performance of the algorithms the normalised mean square error (\textsc{nmse}) and F measure are used. We demonstrate the performance on small datasets to highlight that the proposed algorithm can infer accurate predictions when the dataset size is not sufficient for the \textsc{lista} to learn.

\begin{figure}[t!]
\centering
%\subfloat[\textsc{nmse} on train]{\includegraphics[width=0.5\columnwidth]{graphics/synthetic_number_of_layers/nmse_train}}
%~
\subfloat[\textsc{nmse} for different $L$]{\includegraphics[width=0.4\columnwidth]{graphics/synthetic_number_of_layers/nmse_validation}
\label{fig:nmse_n_layers_synthetic}}
%\subfloat[F measure on train]{\includegraphics[width=0.5\columnwidth]{graphics/synthetic_number_of_layers/f_measure_train}}
%~
\subfloat[F measure for different $L$]{\includegraphics[width=0.4\columnwidth]{graphics/synthetic_number_of_layers/f_measure_validation}
\label{fig:f_meas_n_layers_synthetic}}\\[-13pt]
\subfloat[\textsc{nmse} for different $K$]{\includegraphics[width=0.4\columnwidth]{graphics/synthetic_undersampling/nmse_validation}
\label{fig:nmse_undersampling_synthetic}}
%\subfloat[F measure on train]{\includegraphics[width=0.5\columnwidth]{graphics/synthetic_undersampling/f_measure_train}}
%~
\subfloat[F measure for different $K$]{\includegraphics[width=0.4\columnwidth]{graphics/synthetic_undersampling/f_measure_validation}
\label{fig:f_meas_undesampling_synthetic}}
\caption{Predictive accuracy on the synthetic data: the top row for different numbers of layers (for neural networks) or iterations (for baselines),  the bottom row for different sizes of observations}
\label{fig:number_of_layers_synthetic}
\end{figure}

\subsection{Predictive performance on synthetic data}
First, the predictive performance of the proposed Bayesian \textsc{lista} is analysed on synthetic data. We generate $N_\text{train}=500$ sparse coefficients vectors $\boldsymbol\beta^{(n)}$ each of size $D = 100$. Coefficients $\boldsymbol\beta^{(n)}$ are generated from the spike and slab distribution with truncated slab: each component $[\beta^{(n)}]_{d}$ is zero with the probability $0.8$ or is from the standard Gaussian distribution without interval $(-0.1, 0.1)$ with the probability $0.2$. To simulate sparse observations, we generate the random Gaussian design matrix $\mathbf{X} \in \mathbb{R}^{K \times D}$.  The observations are generated according to (\ref{eq:regression_problem}) with the zero-mean Gaussian noise with the standard deviation $0.5$. The shrinkage parameter is set to~$\lambda = 0.1$.

In Figures \ref{fig:nmse_n_layers_synthetic} and \ref{fig:f_meas_n_layers_synthetic} predictive performance for different number of layers $L$ is presented. The observation size is set to $K=50$. The Bayesian \textsc{lista} outperforms \textsc{lista} in terms of both measures. Although the baselines \textsc{ista} and \textsc{fista} show better performance in terms of F measure, the Bayesian \textsc{lista} has the lowest \textsc{nmse}.

Figures~\ref{fig:nmse_undersampling_synthetic} and~\ref{fig:f_meas_undesampling_synthetic} give the results of predictive performance for different observation sizes $K$. The number of layers is set as $L=4$. In the previous experiment Bayesian and classical \textsc{lista} show similar results with this number of layers. The results of Figures~\ref{fig:nmse_undersampling_synthetic} and~\ref{fig:f_meas_undesampling_synthetic} confirm this competitive behaviour between two \textsc{lista} networks. Baselines show similar results in terms of F measure and underperform in terms of \textsc{nmse}. Overall, the experiments on the synthetic data demonstrate that the Bayesian \textsc{lista} provides competitive results in terms of predictive accuracy.
